Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf = read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
qplot(x = age, y = friend_count, data = pf)

ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_point(alpha = 0.05) +
  xlim(13, 90)

ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_jitter(alpha = 0.05) +
  xlim(13, 90)
  
ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_point(alpha = 0.05, size = 1, position = position_jitter(h=0)) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

names(pf)
  
ggplot(aes(x= age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 0.02, size = 1, position = position_jitter(h=0)) +
  xlim(13, 90) +
  ylim(1, 2000) +
  coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

pf.fc_by_age = pf %>%
  group_by(age) %>% 
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% 
  arrange(age)

ggplot(pf.fc_by_age, aes(x = age)) +
  geom_line(aes(y = friend_count_mean, color = 'red')) +
  geom_line(aes(y = friend_count_median, color = 'blue')) +
  coord_cartesian(ylim=c(0,500))
```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

names(pf)

ggplot(aes(x= age, y = friend_count), data = pf) +
  geom_point(alpha = 0.02, size = 1, position = position_jitter(h=0), color='orange') +
  xlim(13, 90) +
  ylim(0, 2000) +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args=0.9, color='dark red', linetype = 2) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args=0.5, color='dark red') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args=0.1, color='dark red', linetype = 2)


```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)

with(pf, cor.test(age, friend_count))

pf %>%
  subset(age <= 70) %>% 
  with(cor.test(age, friend_count))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(data = pf, aes(x = likes_received, y = www_likes_received)) +
  geom_point(alpha = 0.01) +
  xlim(1, 50000) +
  ylim(1, 50000) +
  coord_trans(x = 'log10', y = 'log10')

ggplot(data = pf, aes(x = likes_received, y = www_likes_received)) +
  geom_point(alpha = 0.01) +
  xlim(0, quantile(pf$likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'blue')

pf %>%
  with(cor.test(x = likes_received, y = www_likes_received))

pf %>%
  subset(likes_received < quantile(likes_received, 0.95)) %>% 
  with(cor.test(x = likes_received, y = www_likes_received))
  
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data("Mitchell")

?Mitchell

Mitchell$Month %% 12

Mitchell %>% 
  ggplot(aes(x = Month, y = Temp)) +
    geom_point()

Mitchell %>% 
  ggplot(aes(x = Month %% 12, y = Temp)) +
    geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
Mitchell %>% 
  with(cor.test(x = Month, y = Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
Mitchell$Month

Mitchell %>% 
  ggplot(aes(x = Month, y = Temp)) +
    geom_point() +
    scale_x_continuous(breaks = seq(0, 203, 12))

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

pf$dob_month

pf %>%
  mutate(age_with_months = age + (1 - dob_month / 12))

```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months = pf %>%
  mutate(age_with_months = age + (1 - dob_month / 12)) %>%
  group_by(age_with_months) %>% 
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% 
  arrange(age_with_months)

pf.fc_by_age_months
```

Programming Assignment
```{r Programming Assignment}

pf.fc_by_age_months %>% 
  subset(age_with_months < 71) %>% 
  ggplot(aes(x = age_with_months, y = friend_count_mean)) +
    geom_line()

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

