Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf = read.csv('pseudo_facebook.tsv', sep='\t')
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)

names(pf)

qplot(x=dob_day, data=pf) +
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(x=dob_day, data=pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count, data=pf,
      binwidth=25)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count, data=pf,
      binwidth=25) +
  scale_x_continuous(limits=c(0, 1000), breaks=seq(0,1000,25))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x=friend_count, data=pf,
      binwidth=25) +
  scale_x_continuous(limits=c(0, 1000), breaks=seq(0,1000,25)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x=friend_count, data=subset(pf, !is.na(gender)),
      binwidth=25) +
  scale_x_continuous(limits=c(0, 1000), breaks=seq(0,1000,25)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x=tenure/365, data=pf,
      binwidth=0.1, color=I('grey'), fill=I('#00AA00')) +
  scale_x_continuous(limits=c(0, 8), breaks=seq(0, 20, 1))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x=tenure/365, data=pf,
      binwidth=1, color=I('grey'), fill=I('#00AA00')) +
  scale_x_continuous(limits=c(0, 8), breaks=seq(0, 20, 1))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

``` {r Transforming Data}
install.packages('gridExtra')
library(gridExtra)

summary(pf$friend_count)


grid.arrange(
  qplot(x=friend_count, data=pf) + scale_x_continuous(limits=c(0,1000)),
  qplot(x=friend_count, data=pf) + scale_x_continuous(trans="log10", limits=c(1,5000)),
  qplot(x=friend_count, data=pf) + scale_x_continuous(trans="sqrt", limits=c(0,3000))
)
```


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
names(pf)

summary(pf$www_likes)

ggplot(aes(x = www_likes, y = ..count../sum(..count..)),
       data=subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color=gender)) +
  scale_x_continuous(trans="log10", limits=c(1,15000))

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)

```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x=www_likes, y=friend_count,
      data=subset(pf, !is.na(gender)),
      geom="boxplot"
      ) +
  scale_y_continuous(trans="log10", limits=c(1,5000))
  facet_wrap(~gender)

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
names(pf)
by(pf$friendships_initiated, pf$gender, summary)

ggplot(aes(x=friendships_initiated, y=..density..),
      data=subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color=gender)) +
  scale_x_continuous(trans="log10")

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
pf$mobile_check_in = factor(ifelse(pf$mobile_likes > 0, 1, 0))

by(pf$mobile_check_in, pf$gender, summary)
x = summary(pf$mobile_check_in)
x[2] / sum(x)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!